Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-19 VMCAI-revised-version/final.tex
effectivity

@@

In this section we study VASS. Since they are ideally-effective and completion-$\post$-effective WSTS, they inherit the decidability results for WSTS in the case 
$\Safe = \uparrow \Safe$. Lacking downward compatibility or a more relaxed hypothesis that
for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 
VASS do not inherit the decidability results for WSTS in the case $\Safe = \downarrow \Safe$.
In this section, we work to re-establish decidability results for VASS when $\Safe$ is downward-closed. We also extend decidability to resilience for semilinear sets rather than simply upward and downward-closed ones.

@@

\begin{proof}
We consider in VASS the case of when $\Safe$ is semilinear
and $\Bad = S \setminus \Safe$ is semilinear too.
We want to use the fact that
it is decidable whether $\post^*(X) \subseteq \pred^*(Y)$ when 
$X$ and $Y$ are both semilinear sets~\cite{DBLP:journals/corr/abs-2207-02697}.
{\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
We have $\post^*(\Bad) \setminus \Safe = \Bad$ by extension of $S \setminus \Safe = \Bad$.
Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
\[\post^*(\Bad) \subseteq \pred^*(\Safe) \quad \text{iff} \quad \Bad \subseteq \pred^*(\Safe).\]
and hence, {\sc Resilience} is decidable. 
The reachability relation of 
%integer $d$-VASS
 $\mathbb{Z}^d-$VASS is definable by a well-formed formula with no free variables in the first-order theory of the integers with addition and orders (Presburger arithmetic) in $\mathbb{Z}^{2d}$, hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets. \qed
\end{proof}

@@

$\mathds{Q}^d$-VASS~\cite{DBLP:journals/tocl/BlondinFHH17} (or continuous VASS) are a relaxation of classical discrete $d$-VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.

@@

The above results concern only unbounded resilience. We consider now 
{\sc Bounded resilience} and {\sc $k$-resilience}, and focus on the length of
the path from $\Bad$ to $\Safe$
when $\Safe$ is downward-closed. Unfortunately, we have the following:
\begin{proposition}
{\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \downarrow \Safe$.
\end{proposition}

>>>>

\alain{ok pour la notation $q(v)$, je vois qu'elle est déjà utilisée}

@@

In this section we study VASS. Since they enjoy many properties of effectivity ($\leq$ is ideally-effective, VASS are $\post$-effective, $\post$-ideal-effective, $\pred$-basis-effective, $\post^*$-basis-effective), they inherit the decidability results for WSTS in the case 
$\Safe = \uparrow \Safe$. Lacking downward compatibility or a more relaxed hypothesis that
for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 
VASS do not inherit the decidability results for WSTS in the case $\Safe = \downarrow \Safe$.
In this section, we work to re-establish decidability results for VASS when $\Safe$ is downward-closed. We also extend decidability to resilience for semilinear sets rather than simply upward and downward-closed ones.

@@

\begin{proof}
We consider the case where $\Safe$ is semilinear
and $\Bad = S \setminus \Safe$ is semilinear too.
We want to use the fact that, for VASS,
it is decidable whether $\post^*(X) \subseteq \pred^*(Y)$ when 
$X$ and $Y$ are both semilinear sets~\cite{DBLP:journals/corr/abs-2207-02697}.
{\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
We have $\post^*(\Bad) \setminus \Safe = \Bad$ by extension of $S \setminus \Safe = \Bad$.
Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
\[\post^*(\Bad) \subseteq \pred^*(\Safe) \quad \text{iff} \quad \Bad \subseteq \pred^*(\Safe).\]
and hence, {\sc Resilience} is decidable. 
The reachability relation of 
%integer $d$-VASS
 $\mathbb{Z}^d-$VASS is definable by a well-formed formula with no free variables in the first-order theory of the integers with addition and orders (Presburger arithmetic) in $\mathbb{Z}^{2d}$, hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets. \qed
\end{proof}

@@

$\mathds{Q}^d$-VASS~\cite{DBLP:journals/tocl/BlondinFHH17} (or continuous VASS) are a relaxation of classical (discrete) $d$-VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.

@@

The above results concern only unbounded resilience. We consider now 
{\sc Bounded resilience} and {\sc $k$-resilience}, and focus on the length of
the path from $\Bad$ to $\Safe$
when $\Safe$ is downward-closed. Surprisingly, we have the following result for VAS (not VASS):
\begin{proposition}
{\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \downarrow \Safe$.
\end{proposition}
